---
title: "A Demonstration Flipbook"
subtitle: "Selecting and Plotting A Variable Across 2 Groups "  
author: 
  - "Peter Higgins"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = FALSE, fig.retina = 3)

library(flipbookr)
library(xaringan)
library(xaringanthemer)
library(tidyverse)
library(medicaldata)
library(janitor)
prostate <- medicaldata::blood_storage %>% clean_names() %>% as_tibble()
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
xaringanthemer::style_duo_accent(
  primary_color = "#00274C",
  secondary_color = "#FFCB05",
  header_font_google = google_font("Lato"),
  code_font_google   = google_font("Roboto Mono"),
  base_font_size = "20px",
  code_font_size = "0.8rem",
)
```

```{r xaringanExtra-share-again, echo=FALSE}
xaringanExtra::use_share_again()
# adds share bar to flipbooks
```

### A Demonstration Flipbook

#### Each Line of Code Produces the Results on the Right

- The pipe **%>%** takes the result of one line of code and 'pipes' it into the next function. <br>It can be read as "and then..."

- The **select** function selects columns in a dataframe.

- The **group_by** function groups the observations(rows) of data into groups

- the **ggplot** function plots data according to a layered model of graphics called the "grammar of graphics".

- each step in the ggplot adds to or changes the plot. 


---

```{r demo1, include = FALSE, eval = FALSE}

prostate %>% 
# select 2 columns
  select(p_vol, recurrence) %>% 
# slow plot 
  ggplot() +
  aes(x = factor(recurrence)) +
  aes(y = p_vol) +
  aes(color = factor(recurrence)) +
  aes(fill = factor(recurrence)) +
  geom_violin(alpha = 0.3) +
  labs(x = "Recurrence", 
       y = "Prostate Volume") +
  theme_linedraw(base_size = 14) +
  theme(legend.position = c(0.5, 0.8)) +
  scale_color_discrete(name = "Recurrence", 
    labels = c("No Recurrence", "Recurrence")) +
  guides(fill = FALSE)
  


```

`r chunk_reveal("demo1", 
    break_type = "auto", 
    title = "Demonstration 1: Select Variables, then Violin Plot")`


