---
title: "Embedded Flipbook"
author: "Peter Higgins"
date: "10/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(flipbookr)
library(tidyverse)
library(janitor)
library(medicaldata)
library(xaringan)
library(xaringanExtra)
data <- medicaldata::blood_storage
```

### Testing how to embed a flipbook in its own Rmd (i.e. a bookdown book)

This is only a test.

```{r wrangle, include = FALSE}
data %>% 
  clean_names() %>% 
  select(recurrence, fam_hx, 
    age, t_stage, units) %>% 
  filter(units > 1) %>% 
  select(-units) %>% 
  arrange(desc(t_stage), 
    desc(recurrence), age)
```

## Data Wranging Example #1

You can:

- Expand the example below to View in full screen by clicking the four arrows icon (Esc to return to small version).
- Go to the next slide or previous slide with the left and right arrow icons/keys. 
- Share on the internet by clicking the Share icon (3 circles connected by 2 lines).

Give it a try.

```{r build_fb1, message=FALSE, warning = FALSE, echo=FALSE}
build_instant_flipbook(chunk_name = "wrangle", use_share_again = T,
      font_size = 110)
```

### Now a plot

This is only a plot.

```{r plot, include = FALSE}
data %>% 
  clean_names() %>% 
  ggplot() +
  aes(x = preop_psa) +
  aes(y = time_to_recurrence) + 
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(~ fam_hx, 
    labeller = 
    labeller(fam_hx=label_both)) + 
  theme_minimal() +
  labs(x = "Preoperative Prostate Specific Antigen (PSA) level") +
  labs(y = "Months to Recurrence of Prostate Cancer") +
  labs(title = "Recurrence of Prostate Cancer by Preop PSA and Family History")
  
```

## Plotting Example #2

You can:

- Expand the example below to View in full screen by clicking the four arrows icon (Esc to return to small version).
- Go to the next slide or previous slide with the left and right arrow icons/keys. 
- Share on the internet by clicking the Share icon (3 circles connected by 2 lines).

Give it a try.

```{r build_fb2, message=FALSE, warning = FALSE, echo=FALSE}
build_instant_flipbook(chunk_name = "plot", 
                use_share_again = T,
                  font_size = 110)
```
