---
title: "How to Use _slice()_ functions to Take Slices of Rows"
subtitle: "Selecting Groups and Extremes of Rows"  
author: 
  - "Peter Higgins"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = FALSE, fig.retina = 3)

library(flipbookr)
library(xaringan)
library(xaringanthemer)
library(tidyverse)
library(medicaldata)
library(lubridate)
library(janitor)
covid_dates <- medicaldata::covid_testing %>% clean_names() %>% as_tibble() %>% 
  mutate(fake_date = ymd("2020-03-01") + pan_day)

```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
xaringanthemer::style_duo_accent(
  primary_color = "#00274C",
  secondary_color = "#FFCB05",
  header_font_google = google_font("Lato"),
  code_font_google   = google_font("Roboto Mono"),
  base_font_size = "20px",
  code_font_size = "0.8rem",
)
```

### How to Use the _slice()_ functions to Take Slices of Rows

#### 

If you want the first 1:N rows, or 50:100 rows, or last 100 rows,
however the dataframe is currently sorted,
you can use the _slice()_ function.

The _slice_max()_ and _slice_min()_ can take rows at top or 
bottom of particular values, using N or proportion.

Example 1: Slice off rows 100:200 from covid_dates.


---

```{r filter1, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(covid_dates) 

covid_dates %>% 
# check if all rows distinct
  slice(100:200)

# see how many rows now

# Format:
#   slice(rownum1 : rownum2)                    <br>
```

`r chunk_reveal("filter1", 
    break_type = "auto", 
    title = "Example 1: Slice Rows 100-200")`

---

```{r filter2, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(covid_dates) 

covid_dates %>% 
  select(age, gender, pan_day) %>% 
# more than 20 because of ties, auto-sorts
  slice_max(age, n=20)

# see how many rows now

# Format:
#   slice_max(variable, n = x)                    <br>
```

`r chunk_reveal("filter2", 
    break_type = "auto", 
    title = "Example 2: Slice Rows with Top 20 age values")`

---

```{r filter3, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(covid_dates) 

covid_dates %>% 
  select(age, gender, pan_day) %>% 
  slice_min(pan_day, prop = 0.01)

# see how many rows now

# Format:
#   slice_min(variable, prop = 0.nn)                    <br>
```

`r chunk_reveal("filter3", 
    break_type = "auto", 
    title = "Example 3: Slice Rows with Bottom 1% of -Earliest- Pandemic Day")`

---

```{r filter4, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(covid_dates) 

covid_dates %>% 
  select(age, gender, pan_day) %>% 
  slice_tail(prop = 0.01)

# see how many rows now

# Format:
#   slice_min(variable, prop = 0.nn)                    <br>
```

`r chunk_reveal("filter4", 
    break_type = "auto", 
    title = "Example 4: Slice Top or Bottom N or %")`


