---
title: "How to Use _filter()_ on Character Variables"
subtitle: "Using _str_detect_ to find the rows you want"  
author: 
  - "Peter Higgins"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = FALSE, fig.retina = 3)

library(flipbookr)
library(xaringan)
library(xaringanthemer)
library(tidyverse)
library(medicaldata)
library(janitor)
cmv <- medicaldata::cytomegalovirus %>% clean_names() %>% as_tibble()
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
xaringanthemer::style_duo_accent(
  primary_color = "#00274C",
  secondary_color = "#FFCB05",
  header_font_google = google_font("Lato"),
  text_font_google   = google_font("Manrope"),
  code_font_google   = google_font("Roboto Mono"),
  base_font_size = "20px",
  code_font_size = "0.8rem",
)
```

### How to Use Filters to Pick Out String Values

#### Leaning on the {stringr} package

Format: <br>

dataset %>% 
  filter(str_detect(variable, "string"))
  
The _str_detect()_ function returns either a TRUR or FALSE for each row in the dataset - a **logical vector**. <br>
The filter function **acts** on this logical vector to filter out rows that are FALSE for your logical statement.
It is a common mistake to rely on str_detect alone, but it won't filter your rows. You need **both** filter and str_detect to get this done.

Let's look at the CMV dataset, which looks at bone marrow transplantation for a variety of cancers. 

Example 1: filter cmv data for cases with a diagnosis that includes the string "oma".

---

```{r filter1, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(cmv) 

cmv %>% 
# selected 3 columns
# see how diagnosis changes w/filter
  select(age, sex, diagnosis) %>% 
  filter(str_detect(diagnosis, "oma"))

# see how many rows now
# check diagnosis
# Format:
# filter(str_detect(variable,"string"))        <br>  


```

`r chunk_reveal("filter1", 
    break_type = "auto", 
    title = "Filter Rows with str_detect on Character Variables")`

---

```{r filter2, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(cmv) 

cmv %>% 
# selected 3 columns
# see how diagnosis changes w/filter
  select(age, race, diagnosis) %>% 
  filter(!str_detect(diagnosis, "Hodgkin"))

# see how many rows now
# check diagnosis
# Format:
# filter(str_detect(variable,"string"))        <br>  


```

`r chunk_reveal("filter2", 
    break_type = "auto", 
    title = "Example 2: Filter Not-Rows (! Negation) with str_detect on Character Variables")`

---

```{r filter3, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(cmv) 

cmv %>% 
# selected 3 columns
# see how diagnosis changes w/filter
  select(sex, race, diagnosis) %>% 
  filter(str_detect(diagnosis, "lympho.+ic"))

# see how many rows now
# check diagnosis
# Format:
# filter(str_detect(variable,"string"))        <br>  


```

`r chunk_reveal("filter3", 
    break_type = "auto", 
    title = "Example 3: Filter Rows with a Little Bit of Regex<br>.+ means any character from 1 to N times")`

---

```{r filter4, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(cmv) 

cmv %>% 
# selected 3 columns
# see how diagnosis changes w/filter
  select(sex, race, diagnosis) %>% 
  filter(str_detect(diagnosis, "myelo*"))

# see how many rows now
# check diagnosis
# Format:
# filter(str_detect(variable,"string"))        <br>  


```

`r chunk_reveal("filter4", 
    break_type = "auto", 
    title = "Example 4: Filter Rows with a Little Bit of Regex<br> * means any character ")`

---

### Where To Learn More about Regex (Regular Expressions)

in R, click: https://stat.ethz.ch/R-manual/R-devel/library/base/html/regex.html 

In general, click here: https://regexone.com/

and/or click here: https://www.sitepoint.com/learn-regex/

and/or click here: https://github.com/ziishaned/learn-regex
