---
title: "How to Use the _filter()_ function for numeric values"
subtitle: "Selecting specific rows"  
author: 
  - "Peter Higgins"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = FALSE, fig.retina = 3)

library(flipbookr)
library(xaringan)
library(xaringanthemer)
library(tidyverse)
library(medicaldata)
library(janitor)
prostate <- medicaldata::blood_storage %>% clean_names() %>% as_tibble()
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
xaringanthemer::style_duo_accent(
  primary_color = "#00274C",
  secondary_color = "#FFCB05",
  header_font_google = google_font("Lato"),
  code_font_google   = google_font("Roboto Mono"),
  base_font_size = "20px",
  code_font_size = "0.8rem",
)
```

### How to Use Filter to Pick Out a Subset of Rows

#### We will Start with Numeric Variables

Format: <br>

`dataset %>% 
  filter(var > value)` # you can use <=, ==, <, >, >=, != etc.

As long as what is in the parentheses evaluates to a logical vector - <br>
a string of TRUE or FALSE values, one for each row in your dataset.

You can Remember that filte**R** is for selecting **R**ows because it ends with an **R**.

Remember that **testing** for equality in a logic test requires 2 = =,
not just one =. This is a **very** common syntax error in R code.

You can also use *near* to get numbers in a range

Example 1: filter prostate data for cases with cancer recurrence.

---

```{r filter1, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(prostate) 

prostate %>% 
# check next line: some recur, some do not
# selected 4 columns
  select(age, aa, fam_hx, recurrence) %>% 
  filter(recurrence == 1) 

# see how many rows x cols now
# 
# Format:
# filter(variable [comparison] value)        <br>  


```

`r chunk_reveal("filter1", 
    break_type = "auto", 
    title = "Example 1: filter prostate data for cases with cancer recurrence == 1")`


---

```{r filter2, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(prostate) 

prostate %>% 
# check next line: some recur, some do not
# selected 4 columns
  select(age, recurrence, preop_psa) %>% 
  filter(preop_psa > 20) 

# see how many rows now
# 
# Format:
# filter(variable [comparison] value)        <br>  


```

`r chunk_reveal("filter2", 
    break_type = "auto", 
    title = "Example 2: Filter Preoperative PSA for Values > 20")`

---

```{r filter3, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(prostate) 

prostate %>% 
# check next line: some recur, some do not
# selected 4 columns
  select(age, recurrence, preop_psa) %>% 
  filter(near(preop_psa, 10)) 

# see how many rows now
# 
# Format:
# filter(variable [comparison] value)        <br>  


```

`r chunk_reveal("filter3", 
    break_type = "auto", 
    title = "Example 3: Filter Preoperative PSA for Values very close to 10")`

---

```{r filter4, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(prostate) 

prostate %>% 
# check next line: some recur, some do not
# selected 4 columns
  select(age, recurrence, preop_psa) %>% 
  filter(near(preop_psa, 17, tol = 1.5)) 

# see how many rows now
# 
# Format:
# filter(variable [comparison] value)        <br>  


```

`r chunk_reveal("filter4", 
    break_type = "auto", 
    title = "Example 4: Filter Preoperative PSA for Values within 1.5 units of 17")`


---

```{r filter5, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(prostate) 

prostate %>% 
# check next line: some recur, some do not
# selected 4 columns
  select(age, recurrence, preop_psa) %>% 
  filter(between(preop_psa, 10, 13)) 

# see how many rows now
# 
# Format:
# filter(variable [comparison] value)        <br>  


```

`r chunk_reveal("filter5", 
    break_type = "auto", 
    title = "Example 5: Filter Preoperative PSA for Values between 10 and 13")`

---

```{r filter5, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(prostate) 

prostate %>% 
# check next line: some recur, some do not
# selected 4 columns
  select(age, recurrence, preop_psa) %>% 
  filter(preop_psa == 10 | preop_psa == 13 |preop_psa == 17)

# see how many rows now
# 
# Format:
# filter(variable [comparison] value)        <br>  


```

`r chunk_reveal("filter5", 
    break_type = "auto", 
    title = "Example 5: Filter Preoperative PSA for Values of 10 or 13 or 17")`
