---
title: "How to Use  _filter()_ functions to Remove Rows with Missing Values"
subtitle: "Selecting specific rows with or without NAs"  
author: 
  - "Peter Higgins"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = FALSE, fig.retina = 3)

library(flipbookr)
library(xaringan)
library(xaringanthemer)
library(tidyverse)
library(medicaldata)
library(lubridate)
library(janitor)
covid_dates <- medicaldata::covid_testing %>% clean_names() %>% as_tibble() %>% 
  mutate(fake_date = ymd("2020-03-01") + pan_day)
bmt_dates <- medicaldata::cytomegalovirus %>% clean_names() %>% as_tibble() %>% 
  mutate(fake_dx_date = ymd("2014-01-01") + days(round(runif(1,1,364))),
         fake_bmt_date = fake_dx_date + days(round(time_to_transplant*30)),
         fake_agvhd_date = fake_bmt_date + days(round(time_to_agvhd*30)))

# add NAs
covid_dates$fake_date[c(7,11,19,33,37,42,66)] <- NA
bmt_dates$fake_dx_date[c(2,8,16,32)] <- NA
bmt_dates[c(2,5), ] <- NA
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
xaringanthemer::style_duo_accent(
  primary_color = "#00274C",
  secondary_color = "#FFCB05",
  header_font_google = google_font("Lato"),
  code_font_google   = google_font("Roboto Mono"),
  base_font_size = "20px",
  code_font_size = "0.8rem",
)
```

```{r xaringanExtra-share-again, echo=FALSE}
xaringanExtra::use_share_again()
# adds share bar to flipbooks
```

### How to Use the _filter()_ function for NA or *not* NA values

#### 

Format: <br>

`dataset %>% 
  filter(is.na(variable)))` 

Negation is done by preceding with ! , <br>
so !is.na(variable) produces a logical vector which is  <br>
TRUE for all _non-missing_ value rows.

**R**emember that filte**R** is for selecting **R**ows because it ends with an **R**.

Let's walk through some **filter** Examples!


---

```{r filter1, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(covid_dates) 

covid_dates %>% 
# selected 3 columns
# see how date changes w/filter
  select(fake_date, result, gender) %>% 
  filter(is.na(fake_date))

# see how many rows now
# check dates - should be NA

# Format:
#   filter(is.na(variable))                 <br>
```

`r chunk_reveal("filter1", 
    break_type = "auto", 
    title = "Example 1/4: Filter Rows for Missing Dates - find which Ones are Missing")`


---

```{r filter2, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(bmt_dates) 

bmt_dates %>% 
# selected 3 columns
# see how date changes w/filter
  select(fake_dx_date, race, donor_sex) %>% 
  filter(!is.na(fake_dx_date))

# see how many rows now
# check dates - should be NOT NA

# Format:
#   filter(!is.na(variable))                 <br>
```

`r chunk_reveal("filter2", 
    break_type = "auto", 
    title = "Example 2/4: Filter Rows for Non-Missing Dates with !is.na()")`

---

```{r filter3, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(bmt_dates) 

bmt_dates %>% 
# selected 3 columns
# remove the 2 empty rows with janitor
  select(fake_dx_date, race, donor_sex) %>% 
  remove_empty("rows")

# see how many rows now
# check dates - should be NOT NA

# Format:
#   remove_empty("rows") (or "cols")                 <br>
```

`r chunk_reveal("filter3", 
    break_type = "auto", 
    title = "Example 3/4: Filter Rows for Non-Missing Dates - when you have (completely) Empty Rows <br>the {janitor} package can help")`

---

```{r filter4, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(bmt_dates) 

bmt_dates %>% 
#   remove incomplete cases with drop_na
  drop_na()

# see how many rows now - fewer!
# now all vars should be NOT NA

# Format:
#   drop_na()                                        <br>                
```

`r chunk_reveal("filter4", 
    break_type = "auto", 
    title = "Example 4/4: Filter Rows for Complete Cases - with zero NAs in that row/observation")`

---
class: inverse, center

# End of This Flipbook

## On to The Coding Exercises!
