---
title: "How to Use Multiple _filter()_ functions with Boolean logic"
subtitle: "Selecting specific rows with multiple criteria"  
author: 
  - "Peter Higgins"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = FALSE, fig.retina = 3)

library(flipbookr)
library(xaringan)
library(xaringanthemer)
library(tidyverse)
library(medicaldata)
library(janitor)
prostate <- medicaldata::blood_storage %>% clean_names() %>% as_tibble()
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
xaringanthemer::style_duo_accent(
  primary_color = "#00274C",
  secondary_color = "#FFCB05",
  header_font_google = google_font("Lato"),
  text_font_google   = google_font("Manrope"),
  code_font_google   = google_font("Roboto Mono"),
  base_font_size = "20px",
  code_font_size = "0.8rem",
)
```

### How to Use Multiple Filters to Pick Out a Subset of Rows

#### Combining Logical Statements

Format: <br>

`dataset %>% 
  filter(var1 > value & var2 > value)` 

You can combine statements with the following:
- & (AND) (age >60 & fam_hx < 1)
- | (OR)  (age >65 | fam_hx == 1)
- ! (NOT) (age >= 65 & fam_hx ! = 1)
- XOR (exclusive OR - only if exactly one of the two is true
filter(xor(age >= 65, fam_hx == 1)

You can **R**emember that filte**R** is for selecting **R**ows because it ends with an **R**.

Example 1: filter prostate data for cases with age >65 and tumor volume >1.


---

```{r filter2, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(prostate) 

prostate %>% 
# selected 3 columns
# see how age and t_vol change w/filter
  select(age, t_vol, recurrence) %>% 
  filter(age > 65 & t_vol>1)

# see how many rows now
# check age and t_vol
# Format:
# filter(variable [comparison] value)        <br>  


```

`r chunk_reveal("filter2", 
    break_type = "auto", 
    title = "Example 2: Filter Rows with Multiple Criteria")`


---

```{r filter3, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(prostate) 

prostate %>% 
# selected 3 columns
# see how age and aa change w/filter
  select(age, aa, t_vol) %>% 
  filter(age > 69 | t_vol > 1) %>% 
  arrange(desc(age))

# see how many rows now
# watch age and aa status
# Format:
# filter(variable [comparison] value)        <br>  


```

`r chunk_reveal("filter3", 
    break_type = "auto", 
    title = "Example 3: Filter Rows with OR Criteria")`

---

```{r filter4, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(prostate) 

prostate %>% 
# selected 3 columns
# see how age and aa change w/filter
  select(age, aa, t_vol) %>% 
  filter(xor(age > 69, t_vol > 1)) %>% 
  arrange(desc(age))

# see how many rows now
# watch age and aa status
# Format:
# filter(variable [comparison] value)        <br>  


```

`r chunk_reveal("filter4", 
    break_type = "auto", 
    title = "Example 4: Filter Rows with XOR Criteria")`


---

```{r filter5, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(prostate) 

prostate %>% 
# selected 3 columns
# see how age and aa change w/filter
  select(age, aa, recurrence) %>% 
  filter(age > 64 & aa != 1) %>% 
  arrange(desc(age))

# see how many rows now
# watch age and aa status
# Format:
# filter(variable [comparison] value)        <br>  


```

`r chunk_reveal("filter5", 
    break_type = "auto", 
    title = "Filter Rows with AND, NOT Criteria")`

