---
title: "How to Use _distinct()_ and _get_dupes()_ to find distinct rows"
subtitle: "Selecting Distinct vs. Duplicated Rows"  
author: 
  - "Peter Higgins"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = FALSE, fig.retina = 3)

library(flipbookr)
library(xaringan)
library(xaringanthemer)
library(tidyverse)
library(medicaldata)
library(lubridate)
library(janitor)
covid_dates <- medicaldata::covid_testing %>% clean_names() %>% as_tibble() %>% 
  mutate(fake_date = ymd("2020-03-01") + pan_day)
bmt_dates <- medicaldata::cytomegalovirus %>% clean_names() %>% as_tibble() %>% 
  mutate(fake_dx_date = ymd("2014-01-01") + days(round(runif(1,1,364))),
         fake_bmt_date = fake_dx_date + days(round(time_to_transplant*30)),
         fake_agvhd_date = fake_bmt_date + days(round(time_to_agvhd*30)))

```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
xaringanthemer::style_duo_accent(
  primary_color = "#00274C",
  secondary_color = "#FFCB05",
  header_font_google = google_font("Lato"),
  code_font_google   = google_font("Roboto Mono"),
  base_font_size = "20px",
  code_font_size = "0.8rem",
)
```

### How to Use the functions to find duplicate or distinct rows

#### 

Imagine that you expect each row to have a unique subject_id, first_name, and last_name.

But you (wisely) decide to check this with the _distinct()_ function. If this nrow is < the original nrow, you have some duplicates.

The {janitor} function _get_dupes()_ can help find these

Example 1: examine the covid_dates data for duplicate entries.


---

```{r filter1, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(covid_dates) 

covid_dates %>% 
# check if all rows distinct
  distinct() %>% 
# check if subject_id is a unique identifier = key
  distinct(subject_id)

# see how many rows now
# subject_id is not a key

# Format:
#   distinct(variable)                    <br>
```

`r chunk_reveal("filter1", 
    break_type = "auto", 
    title = "Example 1: See if Distinct subject_id")`

---

```{r filter2, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(covid_dates) 

covid_dates %>% 
# check if names are distinct
  select(fake_first_name, fake_last_name) %>% 
# check if first & last name is a unique identifier = key
  distinct(fake_first_name, fake_last_name)

# see how many rows now
# fullname is not a key

# Format:
#   distinct(variable1, variable 2)                    <br>
```

`r chunk_reveal("filter2", 
    break_type = "auto", 
    title = "Example 2: See if Distinct by Full Name")`


---

```{r filter3, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(covid_dates) 

covid_dates %>% 
# check if names are distinct
  select(subject_id, fake_first_name, 
         fake_last_name) %>% 
# check if first & last name is a unique identifier = key
  get_dupes(fake_first_name, 
            fake_last_name)

# now you can see the duplicates

# Format:
#   get_dupes(variable1, variable 2)                    <br>
```

`r chunk_reveal("filter3", 
    break_type = "auto", 
    title = "Example 3: Which Rows are Duplicates")`
