---
title: "How to Use _distinct()_ and _get_dupes()_ to find distinct rows"
subtitle: "Selecting Distinct vs. Duplicated Rows"  
author: 
  - "Peter Higgins"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = FALSE, fig.retina = 3)

library(flipbookr)
library(xaringan)
library(xaringanthemer)
library(tidyverse)
library(medicaldata)
library(lubridate)
library(janitor)
covid_dates <- medicaldata::covid_testing %>% clean_names() %>% as_tibble() %>% 
  mutate(fake_date = ymd("2020-03-01") + pan_day)
bmt <- medicaldata::cytomegalovirus %>% 
  clean_names() %>% 
  as_tibble()
bmt_extra <- bmt %>% 
  slice(40:58) 
bmt <-bind_rows(bmt, bmt_extra) 
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
xaringanthemer::style_duo_accent(
  primary_color = "#00274C",
  secondary_color = "#FFCB05",
  header_font_google = google_font("Lato"),
  code_font_google   = google_font("Roboto Mono"),
  base_font_size = "20px",
  code_font_size = "0.8rem",
)
```

```{r xaringanExtra-share-again, echo=FALSE}
xaringanExtra::use_share_again()
# adds share bar to flipbooks
```

### How to Use the distinct() and get_dupes() functions to find duplicate or distinct rows

You can  use _distinct(variable)_ to find all the distinct values of that variable, and therefore how many are duplicates for that variable.

Imagine that you have some duplicate observations in your dataset, due to a problem with data entry, or more likely with a merging of two datasets.

It is important to have a single row for each observation. But it can be hard to find the duplicates in a large dataset.

So you (wisely) decide to check this with the _distinct()_ function. If the resulting number of rows (nrow) is less than the original nrow, you have some duplicates.

The {janitor} function _get_dupes()_ can help find these.


Let's find some duplicates!


---

```{r filter1, include = FALSE, eval = FALSE}
# how many rows when you start

nrow(covid_dates) 

# check if subject_id is a unique identifier 
covid_dates %>% 
  distinct(subject_id)

# see how many rows now
# down to 12,344 - subject_id is not a unique key

# Format:
#   distinct(variable)                    <br>
```

`r chunk_reveal("filter1", 
    break_type = "auto", 
    title = "Example 1/5: Are Some Subjects Being Tested Multiple Times?")`

---

```{r filter2, include = FALSE, eval = FALSE}
# how many rows when you start

nrow(covid_dates) 

covid_dates %>% 
# check if names are distinct
  select(fake_first_name, fake_last_name) %>% 
# check if first & last name is a unique identifier = key
  distinct(fake_first_name, fake_last_name)

# see how many rows now - 8,286
# lots of full names are repeated

# Format:
#   distinct(variable1, variable 2)                    <br>
```

`r chunk_reveal("filter2", 
    break_type = "auto", 
    title = "Example 2/5: See if Distinct by Full Name")`


---

```{r filter3, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(covid_dates) 

covid_dates %>% 
# check if names are distinct
  select(subject_id, fake_first_name, 
         fake_last_name) %>% 
# check if first & last name is unique id
  get_dupes(fake_first_name, 
            fake_last_name)

# now you can see the duplicates

# Format:
#   get_dupes(variable1, variable 2)                    <br>
```

`r chunk_reveal("filter3", 
    break_type = "auto", 
    title = "Example 3/5: Which Rows are Duplicates")`

---

```{r filter4, include = FALSE, eval = FALSE}
# how many rows in bmt

nrow(bmt) 

bmt %>% 
  distinct()

# 64 rows instead of 83 - some duplicates

# Format:
#   distinct(dataset)                    <br>
```

`r chunk_reveal("filter4", 
    break_type = "auto", 
    title = "Example 4/5: Are All Rows in BMT distinct?")`
    
---

```{r filter5, include = FALSE, eval = FALSE}
# how many rows in bmt
nrow(bmt) 

bmt %>% 
  get_dupes()

# 438 duplicate rows

# Format:
#   get_dupes(dataset)                    <br>
```

`r chunk_reveal("filter5", 
    break_type = "auto", 
    title = "Example 5/5: Which Rows are Duplicates")`
    
---
class: inverse, center

# End of This Flipbook

## On to The Coding Exercises!
