---
title: "How to Use the _filter()_ function for numeric values"
subtitle: "selectinf specific rows"  
author: 
  - "Peter Higgins"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = FALSE, fig.retina = 3)

library(flipbookr)
library(xaringan)
library(xaringanthemer)
library(tidyverse)
library(medicaldata)
library(janitor)
prostate <- medicaldata::blood_storage %>% clean_names() %>% as_tibble()
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF"
)
```

### How to Use Filter to Pick Out a Subset of Rows

#### We will Start with Numeric Variables

Format: <br>

`dataset %>% 
  filter(var > value) # you can use <=, ==, <, >, >=, etc.`

As long as what is in the parentheses evaluates to a logical vector - a string of TRUE or FALSE values, one for each row in your dataset.

You can Remember that filteR is for selecting Rows because it ends with an R.

Example 1: filter `prostate` data for recurrence == 1


---

```{r filter1, include = FALSE, eval = FALSE}
nrow(prostate) # how many rows when you start

prostate %>%
# select the variables needed
  filter(recurrence == 1)) %>% 

# see how many rows now
# 
# Format:
# filter(var [comparison] value)
```

`r chunk_reveal("filter1", 
    break_type = "auto", 
    title = "Filter Rows on Numeric Values")`


