---
title: "How to Use  _filter()_ functions with Dates"
subtitle: "Selecting specific rows based on Dates and Date Ranges"  
author: 
  - "Peter Higgins"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = FALSE, fig.retina = 3)

library(flipbookr)
library(xaringan)
library(xaringanthemer)
library(tidyverse)
library(medicaldata)
library(lubridate)
library(janitor)
covid_dates <- medicaldata::covid_testing %>% clean_names() %>% as_tibble() %>% 
  mutate(fake_date = ymd("2020-03-01") + pan_day)
bmt_dates <- medicaldata::cytomegalovirus %>% clean_names() %>% as_tibble() %>% 
  mutate(fake_dx_date = ymd("2014-01-01") + days(round(runif(1,1,364))),
         fake_bmt_date = fake_dx_date + days(round(time_to_transplant*30)),
         fake_agvhd_date = fake_bmt_date + days(round(time_to_agvhd*30)))
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
xaringanthemer::style_duo_accent(
  primary_color = "#00274C",
  secondary_color = "#FFCB05",
  header_font_google = google_font("Lato"),
  code_font_google   = google_font("Roboto Mono"),
  base_font_size = "20px",
  code_font_size = "0.8rem",
)
```

```{r xaringanExtra-share-again, echo=FALSE}
xaringanExtra::use_share_again()
# adds share bar to flipbooks
```

### How to Use the _filter()_ function for Rows by Date

#### 

Format: <br>

`dataset %>% 
  filter(between(variable, start, end))` 

Note that _between_ is **nested** within filter.

Start and end dates usually use the {lubridate} package, and the format     
`ymd("2015-12-31")`

**R**emember that filte**R** is for selecting **R**ows because it ends with an **R**.

Let's step through some _**filter**_ Examples!


---

```{r filter1, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(covid_dates) 

covid_dates %>% 
# selected 3 columns
# see how date changes w/filter
  select(fake_date, result, gender) %>% 
  filter(between(fake_date, 
      ymd("2020-03-01"), ymd("2020-03-31")))

# see how many rows now
# check dates - between is inclusive
# of end dates - tests on 3/1
# and on 3/31 are included.
# Format:
# filter(between(variable, date, date))       <br>  


```

`r chunk_reveal("filter1", 
    break_type = "auto", 
    title = "Example 1/4: Filter Rows Between Dates")`

---

```{r filter2, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(bmt_dates) 

bmt_dates %>% 
# selected 3 columns
# see how date changes w/filter
  select(fake_bmt_date, sex, age) %>% 
  filter(between(fake_bmt_date, 
      ymd("2008-01-01"), ymd("2018-12-31"))) %>% 
  filter(between(age, 40, 70)) %>% 
  arrange(fake_bmt_date) # sort

# see how many rows now
# sequential filters combine with AND
# check dates 
# Format:
# filter(between(variable, start, end))       <br>  


```

`r chunk_reveal("filter2", 
    break_type = "auto", 
    title = "Example 2/4: Find Bone Marrow Transplants in 2008-2018 in 40-70 yo")`

---

```{r filter3, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(bmt_dates) 

bmt_dates %>% 
# selected 3 columns
# see how date changes w/filter
  select(fake_agvhd_date, sex, age) %>% 
  filter(fake_agvhd_date > 
           today() - months(24)) %>% 
  arrange(fake_agvhd_date) # sort

# see how many rows now
# note we have some "future" fake dates
# Format:
# filter(between(variable, start, end))       <br>  


```

`r chunk_reveal("filter3", 
    break_type = "auto", 
    title = "Example 3/4: Find Acute GVHD Cases in last 24 Months")`

---

```{r filter4, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(bmt_dates) 

bmt_dates %>% 
# selected 3 columns
# see how date changes w/filter
  select(fake_bmt_date, sex, age) %>% 
# wday gives weekday 1-7, labeled
# note one NA date
  mutate(weekday = wday(fake_bmt_date, 
              label=TRUE)) %>%
# %in% must be in a vector of values
  filter(wday(fake_bmt_date) %in% c(1,7)) %>% 
  arrange(fake_bmt_date) # sort

# see how many rows now
# lots of handy functions in lubridate

```

`r chunk_reveal("filter4", 
    break_type = "auto", 
    title = "Example 4/4: Find BMTs on (Fake) Weekends")`

---
class: inverse, center

# End of This Flipbook

## On to The Coding Exercises!
