---
title: "How to Use _sample()_ functions to Take Random Subsets of Rows"
subtitle: "Selecting Rows with _slice_sample()"  
author: 
  - "Peter Higgins"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = FALSE, fig.retina = 3)

library(flipbookr)
library(xaringan)
library(xaringanthemer)
library(tidyverse)
library(medicaldata)
library(lubridate)
library(janitor)
covid_dates <- medicaldata::covid_testing %>% clean_names() %>% as_tibble() %>% 
  mutate(fake_date = ymd("2020-03-01") + pan_day)

```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
xaringanthemer::style_duo_accent(
  primary_color = "#00274C",
  secondary_color = "#FFCB05",
  header_font_google = google_font("Lato"),
  code_font_google   = google_font("Roboto Mono"),
  base_font_size = "20px",
  code_font_size = "0.8rem",
)
```

### How to Use the _slice()_ functions to Take Slices of Rows

#### 

If you have a very large dataset, and want to develop code on a smaller (but random) sample,
_slice_sample()_ can help.

This is also helpful for sampling for training and testing sets when modeling.

_slice_sample()_ can take n or proportion (prop) arguments

Example 1: Take a random 30% sample of rows from covid_dates.


---

```{r filter1, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(covid_dates) 

covid_dates %>% 
  slice_sample(prop = 0.3)

# see how many rows now

# Format:
#   slice_sample(prop = 0.nn)                    <br>
```

`r chunk_reveal("filter1", 
    break_type = "auto", 
    title = "Example 1: Take a Random 30% Sample of Rows from covid_dates.")`

---

```{r filter2, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(covid_dates) 

# make training set
covid_dates %>% 
  slice_sample(prop = 0.7) ->
training_covid_dates

# now make testing set
setdiff(covid_dates, 
        training_covid_dates) 
# see how many rows in each

# Format:
#   slice_sample(prop = 0.nn)  %>%          <br>
#   setdiff(full_data, .)                  <br>
```

`r chunk_reveal("filter2", 
    break_type = "auto", 
    title = "Example 2: Take a Random 70% Sample for Training and a Complementary 30% for Testing.")`

---

```{r filter3, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(covid_dates) 

covid_dates %>% 
  slice_sample(n = 50)

# see how many rows now

# Format:
#   slice_sample(n = NN)                    <br>
```

`r chunk_reveal("filter3", 
    break_type = "auto", 
    title = "Example 3: Take a Random Sample of 50 Rows from covid_dates.")`

