---
title: "Demo with Tabsets (Panelsets)"
subtitle: "Selecting Rows with _slice_sample()"  
author: 
  - "Peter Higgins"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = FALSE, fig.retina = 3)

library(flipbookr)
library(xaringan)
library(xaringanthemer)
library(tidyverse)
library(medicaldata)
library(lubridate)
library(janitor)
covid_dates <- medicaldata::covid_testing %>% clean_names() %>% as_tibble() %>% 
  mutate(fake_date = ymd("2020-03-01") + pan_day)

```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
xaringanthemer::style_duo_accent(
  primary_color = "#00274C",
  secondary_color = "#FFCB05",
  header_font_google = google_font("Lato"),
  code_font_google   = google_font("Roboto Mono"),
  base_font_size = "20px",
  code_font_size = "0.8rem",
)
```

```{r xaringanExtra-share-again, echo=FALSE}
xaringanExtra::use_share_again()
# adds share bar to flipbooks
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
# gives you tabset option
```
### How to Use the _slice()_ functions to Take Slices of Rows

#### 

If you have a very large dataset, and want to develop code on a smaller (but random) sample,
_slice_sample()_ can help.

This is also helpful for sampling for training and testing sets when modeling.

_slice_sample()_ can take n or proportion (prop) arguments

Let's see some **sampling** examples!


---
.panelset[
.panel[.panel-name[R Code]
```{r filter1, results='hide'}
# how many rows when you start
nrow(covid_dates) 

covid_dates %>% 
  slice_sample(prop = 0.3)

# see how many rows now

# Format:
#   slice_sample(prop = 0.nn) 
```
]

.panel[.panel-name[Results]

```{r , echo=FALSE}
# how many rows when you start
nrow(covid_dates) 

covid_dates %>% 
  slice_sample(prop = 0.3)

# see how many rows now

# Format:
#   slice_sample(prop = 0.nn)  
```
 ]
] 

---

```{r filter2, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(covid_dates) 

# make training set
covid_dates %>% 
  slice_sample(prop = 0.7) ->
training_covid_dates

# now make testing set
covid_dates %>% 
  anti_join(training_covid_dates)->
testing_covid_dates
# see how many rows in each
training_covid_dates
testing_covid_dates
# Format:
#   slice_sample(prop = 0.nn)  
#   set1 %>% anti_join(set2)
```

`r chunk_reveal("filter2", 
    break_type = "auto", 
    title = "Example 2/3: Take a Random 70% Sample for Training and a Complementary 30% for Testing.")`

---

```{r filter3, include = FALSE, eval = FALSE}
# how many rows when you start
nrow(covid_dates) 

covid_dates %>% 
  slice_sample(n = 50)

# see how many rows now

# Format:
#   slice_sample(n = NN)                    <br>
```

`r chunk_reveal("filter3", 
    break_type = "auto", 
    title = "Example 3/3: Take a Random Sample of 50 Rows from covid_dates.")`

---
class: inverse, center

# End of This Flipbook

## On to The Coding Exercises!
