---
title: "Customizing Plot Scales"
subtitle: "Customizing Your Scales in ggplot2"  
author: 
  - "Peter Higgins"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = FALSE, fig.retina = 3)
library(flipbookr)
library(xaringan)
library(xaringanthemer)
library(tidyverse)
library(medicaldata)
library(janitor)
prostate <- medicaldata::blood_storage %>% clean_names() %>% as_tibble()
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
xaringanthemer::style_duo_accent(
  primary_color = "#00274C",
  secondary_color = "#FFCB05",
  header_font_google = google_font("Lato"),
  code_font_google   = google_font("Roboto Mono"),
  base_font_size = "20px",
  code_font_size = "0.8rem",
)
```

```{r xaringanExtra-share-again, echo=FALSE}
xaringanExtra::use_share_again()
# adds share bar to flipbooks
```

### Customizing Plot Scales

#### There are Multiple Functions to Customize the Default Scales of Your ggplot2 plot.

- The {ggplot2} package supplies default scales and labels for your plots so that you can get a reasonable result quickly.

- However, you can customize these scales by adding (+) lines to your plot code that have the format of scales_(x|y|color|shape)_description().

- Let's see an example


---

```{r scales1, include = FALSE, eval = FALSE}

prostate %>% 
# slow plot 
  ggplot() +
  aes(x = factor(recurrence)) +
  aes(y = p_vol) +
  geom_jitter() +
  aes(color = factor(recurrence)) +
  # expand x margins
  scale_x_discrete(expand = expansion(mult=1.2)) +
  # set y limits
  scale_y_continuous(limits= c(0,90)) +
  # format color legend
  scale_color_discrete(name = "Recurrence", 
    labels = c("No Recurrence", "Recurrence")) 
  


```

`r chunk_reveal("scales1", 
    break_type = "auto", 
    title = "Scales 1: x,y, and color")`






---
class: inverse, center

# End of This Flipbook

## Let's Learn How to Do This!

